Regression Discontinuity Design to Measure Incumbency Advantage in US Congressional Elections

### 2023-04-25 paper.qmd
start results

---
title: "Regression Discontinuity Design to Measure Incumbency Advantage in US Congressional Elections"
subtitle: "Final Project for STAT 530"
author: "Tyler Bagwell and Prayag Gordy"
date: "April 24, 2023"
abstract: "The incumbency effect in U.S. elections is a subject of decades of political science research. Why is it, political scientists ask, that legislative and executive bodies with such low overall approval ratings have the vast majority of its members returning term after term? In this paper, we will use a regression discontinuity design to examine the effects of party incumbency in tightly contested congressional districts. Specifically, we examine the effect of the Democratic margin in an election on the Democratic margin in the following election. We found a local average treatment effect of 17.3 percentage points. Our results further support the literature claiming a large incumbency effect, though we show its strength may be decreasing."
format: pdf
toc: true
number-sections: true
execute:
  echo: false
---

@@

>>>>

---
title: "Regression Discontinuity Design to Measure Incumbency Advantage in US Congressional Elections"
subtitle: "Final Project for STAT 530"
author: "Tyler Bagwell and Prayag Gordy"
date: "April 24, 2023"
abstract: "The incumbency effect in U.S. elections is a subject of decades of political science research. Why is it, political scientists ask, that legislative and executive bodies with such low overall approval ratings have the vast majority of its members returning term after term? In this paper, we will use a regression discontinuity design to examine the effects of party incumbency in tightly contested congressional districts. Specifically, we examine the effect of the Democratic margin in an election on the Democratic margin in the following election. We found a local average treatment effect of 17.3 percentage points. Our results further support the literature claiming a large incumbency effect, though we show its strength may be decreasing."
format: pdf
toc: true
number-sections: true
execute:
  echo: false
  message: false
  warning: false
---

```{r message=FALSE, warning=FALSE}
source("main.R")
```

@@

```{r}
#| fig-cap: McCrary density test and estimation for all $t-1$ Democratic vote margins.

lag_any_candidates(DF) |>
  dplyr::pull(prev_dem_margin) |>
  rdd::DCdensity()
```

Second, we assume that the only relevant pretreatment covariates of concern for congressional district-level election outcomes are rates of unemployment, racial demographics, and level of education of a district's citizens. Undoubtedly, this is a gross oversimplification. The number of variables that affect the outcomes of close elections are innumerous, including campaign spending to the election day weather and everything in between, but collecting and analyzing all these data would be out of scope for this class project. Therefore, for the purposes of this small study and its given time constraints, we only consider the three aforementioned covariates.

With these three pretreatment covariates chosen, we estimate the discontinuities in their conditional densities conditional on the $t-1$ Democratic election margin at the zero margin value using local linear regression and find that there are no statistically significant discontinuities in either three covariates’ densities. This result can also be expected just by visual inspection of the covariate value versus the $t-1$ Democratic election margin.
